Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-01-31 TACAS2018/paper.tex
wording

We perform digital control synthesis over a hybrid model, where the plant
exhibits continuous behavior whereas the controller operates in discrete
time and over a quantized domain.  Our model evaluates the effects of the
quantizers (A/D and D/A converters) and of time discretization, as well as
representation errors introduced by modeling artifacts such as an
observer~\cite{astrom1997computer} in a finite precision domain, while
reasoning about high level properties such as stability, robustness, safety,
and response time, which define the behavior of the system.  In particular,
safety requirements are frequently overlooked in conventional feedback
control synthesis, but play an important role in systems engineering.

@@

We use a CounterExample-Guided Inductive Synthesis (CEGIS)~\cite{jha-icse10,
DBLP:conf/pldi/Solar-LezamaRBE05} approach, which we enhance with
abstraction refinement and optimization resulting in the novel CEGIS-OR
(CEGIS with Optimization and Refinement).  This design uses an iterative
process where each iteration performs inductive generalization based on
counterexamples provided by a verification oracle.  Essentially, the
inductive generalization uses information about a limited number of inputs
to make claims in the form of candidate solutions about all the possible
inputs.  By using abstraction refinement, CEGIS-OR allows us to handle
complex systems which are initially represented by coarse abstractions that
only get refined on demand.  Consequently, solutions that do not require a
high level of precision are obtained fast.  Since the use of coarse
abstractions may result in low-quality candidate solutions (i.e.  candidates
that fail the specification on a very large number of inputs), after each
inductive generalization phase we use optimization techniques to improve the
produced candidates.  In particular, our optimization procedure aims at
reducing the size of the reach space of the closed loop system, while
maintaining other dynamical properties of interest (safety, stability, and
response time).

@@

We provide experimental results showing that our approach is able to
efficiently synthesize safe controllers for a set of physical
plant models taken from the digital control literature.

In summary, this paper makes the following original contributions.
%
\begin{enumerate}
\item We develop a model for inductive synthesis that combines CEGIS with
  abstraction refinement and optimization in order to handle complex
  systems. 
%
\item We automatically generate \emph{correct-by-construction} digital
  state-feedback observer controllers that guarantee a given safety property
  alongside stability and response time specifications.  We note that
  observers are much more complex and difficult to synthesize than standard
  full-state feedback controllers (hence the need for abstraction refinement).
%
\item  Our synthesis addresses challenges created by digitalization errors
 such as quantization, and time discretization as well as delays in the
 sampling and processing times which introduce control errors in real
 implementations.
 %
 \item We present a model for soundly evaluating reachability in a combined 
 continuous-discrete time/space. 
% Sound tools for continuous time reachability are scarce and to the authors'
% knowledge no tool provides a model that combines both. 
\end{enumerate}

@@

Program synthesizers are an ideal fit for the synthesis of digital
controllers, since the semantics of programs capture the effects of
finite-precision arithmetic precisely. 
In~\cite{DBLP:conf/cdc/RavanbakhshS15}, the authors use CEGIS for the
synthesis of switching controllers for stabilizing continuous-time plants
with polynomial dynamics.  The work extends to affine systems, but is
limited by the capacity of the state-of-the-art SMT solvers for solving
linear arithmetic.  Since this approach uses switching models instead of
linear dynamics for the digital controller, it avoids problems related to
finite precision arithmetic, but potentially suffers from state-space
explosion.  Moreover, in \cite{DBLP:conf/emsoft/RavanbakhshS16} the same
authors use a CEGIS-based approach for synthesizing continuous-time
switching controllers that guarantee \emph{reach-while-stay} properties of
closed-loop systems, i.e., properties that specify a set of goal states and
safe states (constrained reachability).  This solution is based on
synthesizing control Lyapunov functions for switched systems that yield
switching controllers with a guaranteed minimum dwell time in each mode. 
However, both approaches are unsuitable for the kind of control we seek to
synthesize.

The work in~\cite{hscc-paper} synthesizes stabilizing controllers for
continuous plants
%given as transfer functions
by exploiting bit-accurate verification of software-implemented
digital controllers~\cite{Bessa16} but it is restricted to closed-loop
systems given as transfer function models.

@@

The tool Pessoa~\cite{mazo2010pessoa} synthesizes correct-by-design embedded
control software in a Matlab toolbox.  It is based on the abstraction of a
physical system to an equivalent finite-state machine and on the
verification of reachability properties on it.  Based on this safety
specification, \mbox{Pessoa} can synthesize embedded controller software for
a range of properties.  The embedded controller software can be more
complicated than the state-feedback control we synthesize, and the
properties available cover more detail.  However, relying on state-space
discretization \mbox{Pessoa} is likely to incur scalability limitations. 
Along this research line, \cite{Anta2010,liu16} studies the synthesis of
digital controllers for continuous dynamics, and \cite{zamani2014} extends
the approach to the recent setup of Network Control Systems.

@@

\noindent\textbf{Discretization Effects --}
The classical approach to control synthesis has often disregarded
digitalization effects, whereas more recently modern techniques have focused
on different aspects of discretization, including delayed
response~\cite{Duggirala2015} and Finite Word Length (FWL) semantics, with
the goal either to verify (e.g.,~\cite{daes20161}) or to optimize
(e.g.,~\cite{oudjida2014design}) given implementations.

There are two different problems that arise from FWL semantics.  The first
is the error in the dynamics caused by the inability to represent the exact
state of the physical system, while the second relates to rounding and
saturation errors during computation.  In~\cite{fialho1994stability}, a
stability measure based on the error of the digital dynamics ensures that
the deviation introduced by FWL does not make the digital system unstable. 
A~more recent approach~\cite{DBLP:journals/automatica/WuLCC09} uses
$\mu$-calculus to directly model the digital controller so that the selected
parameters are stable by design.  The analyses
in~\cite{DBLP:conf/hybrid/RouxJG15, DBLP:conf/hybrid/WangGRJF16} rely on an
invariant computation on the discrete system dynamics using Semi-Definite
Programming (SDP).  While the former uses bounded-input and bounded-output
(BIBO) properties to determine stability, the latter uses Lyapunov-based
quadratic invariants.  In both cases, the SDP solver uses floating-point
arithmetic and soundness is checked by bounding the error.  An alternative
is~\cite{park2016scalable}, where the verification of given control code is
performed against a known model by extracting an LTI model of the code by
symbolic execution: to account for rounding errors, an upper bound is
introduced in the verification phase.  The work in
\cite{picasso2003stabilization} introduces invariant sets as a mechanism to
bound the quantization error effect on stabilization as an invariant set
that always converges toward the controllable set.  Similarly,
\cite{liberzon2003hybrid} evaluates the quantization error dynamics and
bounds its trajectory to a known region over a finite time period.  This
technique works for both linear and non-linear systems.

@@

We are interested in control systems with an output $\vec{y}$ that is a linear combination of its states and inputs\footnote{Notice that since we can only measure $\vec{y}$, the state-space $\vec{x}$ is not fully observable.}.
%, which may restrict the observability of the state-space from the output space.
\begin{equation}
\label{eq:ode2}  
\vec{y}(t)=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t), \vec{y} \in \mathbb{R}^o, \mat{C} \in \mathbb{R}^{o \times p}, \mat{D} \in \mathbb{R}^{o \times q}. 
\end{equation}
%
Equations~\eqref{eq:ode1}-\eqref{eq:ode2} are soundly discretized in
time~\cite{middleton1990digital,van1978computing} at constant intervals,
whose duration we refer to as the sample time $T_s$, into the difference
equation
%
\begin{align}
\label{eq:plant}
\left\{
\begin{array}{rl}
\vec{x}_{k+1} =& \mat{A}_d \vec{x}_k+ \mat{B}_d \vec{u}_k\\
\vec{y}_{k} =& \mat{C} \vec{x}_k + \mat{D} \vec{u}_k, 
\end{array}
\right.
\text{ where }
\left\{
\begin{array}{l}
\mat{A}_d=e^{\mat{A}T_s}\\
\mat{B}_d = \int_{t = 0}^{T_s} e^{\mat{A} t} dt\ \mat{B}
\end{array}
\right.,
\end{align} 
%
%where $\mat{A}_d=e^\mat{A}$, $\mat{B}_d = \int_{t = 0}^{T_s} e^{\mat{A} t} dt\ \mat{B}$,
$k \in \mathbb N$ is a discrete counter and $\vec{x}_{0}=\vec{x}(0)$ is the initial state. 
In this work we raise the standard assumption that $\mat{D}=\mat{0}$ \cite{Astrom08}.  
%Also, since $\vec{y}$ can be directly derived from $\vec{x}$, we will omit discussing it in the following unless strictly necessary.

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Semantics of LTI models}\label{sec:model_semantics}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%
The traces of model \eqref{eq:plant}, 
starting from an initial set $X_0\subseteq \mathbb{R}^p$, 
with inputs taken from the set $U \subseteq \mathbb{R}^q$, are sequences 
$ \vec{x}_0 \xrightarrow{\vec{u}_0} \vec{x}_1 \xrightarrow{\vec{u}_1} \vec{x}_2 \xrightarrow{\vec{u}_2} \ldots $, 
%
where $ \vec{x}_0 \in X_0$ and $\forall k\geq 0, \vec{u}_k \in U \text{ and } \vec{x}_{k+1} = \tau(\vec{x}_k,\vec{u}_k) $, 
where 
%
\begin{equation}\label{equ:reachtraj}
\tau(\vec{x}_k,\vec{u}_k) = %\big\{ 
\mat{A}\vec{x}_k + \mat{B}\vec{u}_k
 %\mid \mat{G}\vec{x}_k \leq \vec{h}\big\}\;. 
\end{equation}
%\note{do we actually need to mention the safety spec here given that it is formally introduced only in Sec~\ref{sec:cof_safety}?}
%
%Here $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe a safe set, within the state-space $\mathbb{R}^p$.
%
We generalize this notation to convex sets of initial conditions and inputs ($X_0$ and $U$, respectively), 
and write $\tau(X_0,U)$ to denote the set of states $\{\vec{x} \mid \vec{x} = \tau(\vec{x}_0,\vec{u}),\, \vec{x}_0 \in
X_0,\, \vec{u} \in U \}$
reached from $X_0$ by $\tau$ in one step. 
%
We furthermore define the $n$-reach set $\tau^n(X_0,U)$ as the set of states reached from
$X_0$ via $\tau$ in $n$ steps.
%
\begin{align}\label{equ:reachset}
\tau^n(X_0,U) = \{\vec{x}_n \mid & 
\forall k\in [0,n-1] \, \vec{x}_{k+1}=\tau(\vec{x}_{k},\vec{u}_{k}), \,
\vec{u}_{k} \in U, \,   
\vec{x}_0 \in X_0
\} \;. 
\end{align}
%
Since the transformations $\mat{A}$ and $\mat{B}$ are linear and linear transformations preserve convexity~\cite{boyd2004convex}, 
the sets $X_n = \tau^n(X_0,U)$ are also convex.  
We define the \emph{$n$-reach tube} as 
\begin{equation}\label{equ:reachtube}
\hat{X}_n=\hat{\tau}^n(X_0,U)=\bigcup_{k\in[0,n]} \tau^k(X_0,U)
\end{equation}
as the union of $k$-reach sets over $n$ iterations. 
%
Finally, the set $\hat{X} =\bigcup_{n\geq 0} \tau^n(X_0,U)$, which we refer to as
the reach space of the system, generalizes the previous notions over an unbounded
time horizon (transitive closure).

@@

A closed-loop digital control interconnection with observer is shown in
Figure~\ref{fig:observersystem}.  The linear feedback controller (bottom
left) is characterized by a gain matrix $\mat{K}_{\langle C \rangle}$ that
influences the closed-loop dynamics, and is driven by the observer state
$\hat{\vec{x}}$.  Moreover, $\vec{y}_e=\hat{\vec{y}} - \vec{y}$ is the error
on the output estimator,
%  defined as the difference between the real output $\vec{y}_k$ and estimated one $\hat{\vec{y}_k}$, 
which is amplified by a gain matrix $\mat{L}_{\langle C \rangle}$.  Note
that the controller and the observer are implemented in a discrete domain
(quantities within this domain are identified by the index $\langle C
\rangle$), whereas the plant evolves in a continuous domain.  We further
model the effects of using a discrete controller as the noise signals
$\nu_1$(ADC quantization), $\nu_2$ (DAC quantization), $\nu_1'$ (rounding of
the observer state error), $\nu_2'$ (rounding of the observer state), and
$\nu_t$ (output delay of the digital system) Lastly, $k_r$ is a reference
gain required by the model.  Our goal in this work is to synthesize both
$\mat{K}_{\langle C \rangle}$ and $\mat{L}_{\langle C \rangle}$, so that
specific closed-loop properties of the overall interconnection are
satisfied.

@@

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \section{Acceleration of continuous-time models}\label{sec:continuous_time_accel}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++ 
 The time discretization in Sec.~\ref{sec:model} only 
 allows us to evaluate the state of the plant at integer multiples of the sample time.
 However, in order to synthesize $\mat{K}=\mat{K}_{\langle C \rangle}$ and $\mat{L}=\mat{L}_{\langle C \rangle}$, 
 we must be able to ensure that certain properties of the plant hold at any given real time.
 In order to do so, 
 we accelerate the dynamics of the plant according to the model semantics in Sec.~\ref{sec:model_semantics} and compute the state $\vec{x}_T$  at a given time $T$, 
% namely we find an equation for $\tau^n(\vec{x},\vec{u})$ given $\tau(\vec{x},\vec{u})$. 
namely the state of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time $t=T$.

%\note{ [add?] In this section we work with fully observed models, namely where $y = x$. } 
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \subsection{Accelerated\! dynamics\! for\! continuous-time\! models\! without\! inputs}\label{sec:cont_acc_no_inputs}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 Given a specific time $T$, the state $\vec{x}(t=T)$ can be expressed as a linear function of $\vec{x}_0$ as shown below.
We assume that the system has no inputs, so its dynamics are described by $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$.

>>>>

We perform digital control synthesis over a hybrid model, where the plant
exhibits continuous behavior whereas the controller operates in discrete
time and over a quantized domain.  Our model incorporates the effects of the
quantizers (A/D and D/A converters) and of time discretization, as well as
representation errors introduced by modeling artifacts such as an
observer~\cite{astrom1997computer} in a finite-precision domain, while
reasoning about high-level properties such as stability, robustness, safety,
and response time, which define the behavior of the system.  In particular,
safety requirements are frequently overlooked in conventional feedback
control synthesis, but play an important role in systems engineering.

@@

We use a CounterExample-Guided Inductive Synthesis (CEGIS)~\cite{jha-icse10,
DBLP:conf/pldi/Solar-LezamaRBE05} approach, which we enhance with
abstraction refinement and optimization resulting in the novel CEGIS-OR
(CEGIS with Optimization and Refinement).  This design uses an iterative
process where each iteration performs inductive generalization based on
counterexamples provided by a verification oracle.  Essentially, the
inductive generalization uses information about a limited number of inputs
to make claims in the form of candidate solutions about all the possible
inputs.  By using abstraction refinement, CEGIS-OR allows us to handle
complex systems, which are initially represented by coarse abstractions that
only get refined on demand.  Consequently, solutions that do not require a
high level of precision are obtained fast.  Since the use of coarse
abstractions may result in low-quality candidate solutions (i.e., candidates
that fail the specification on a very large number of inputs), we use
optimization techniques after each inductive generalization phase to improve
the produced candidates.  In particular, our optimization procedure reduces
the size of the reach space of the closed loop system, while maintaining
other dynamical properties of interest (safety, stability, and response
time).

@@

We provide experimental results that show that our approach is able to
efficiently synthesize safe controllers for a set of physical plant models
taken from the digital control literature.

In summary, this paper makes the following original contributions.
%
\begin{enumerate}
%
\item We develop a model for inductive synthesis that combines CEGIS with
  abstraction refinement and optimization in order to handle complex
  systems. 
%
\item We automatically generate \emph{correct-by-construction} digital
  state-feedback observer controllers that guarantee a given safety property
  alongside stability and response time specifications.  We note that
  observers are much more complex and difficult to synthesize than standard
  full-state feedback controllers (hence the need for abstraction refinement).
%
\item  Our synthesis addresses challenges created by digitalization errors
 such as quantization, and time discretization as well as delays in the
 sampling and processing times which introduce control errors in real
 implementations.
 %
 \item We present a model for soundly evaluating reachability in a combined 
 continuous-discrete time/space. 
%
% Sound tools for continuous time reachability are scarce and to the authors'
% knowledge no tool provides a model that combines both. 
\end{enumerate}

@@

Program synthesizers are an ideal fit for the synthesis of digital
controllers, since the semantics of programs capture the effects of
finite-precision arithmetic precisely. 
In~\cite{DBLP:conf/cdc/RavanbakhshS15}, the authors use CEGIS for the
synthesis of switching controllers for stabilizing continuous-time plants
with polynomial dynamics.  The work extends to affine systems, but is
limited by the capacity of the state-of-the-art SMT solvers for solving
linear arithmetic.  Since this approach uses switching models instead of
linear dynamics for the digital controller, it avoids problems related to
finite-precision arithmetic, but potentially suffers from state-space
explosion.  Moreover, in~\cite{DBLP:conf/emsoft/RavanbakhshS16} the same
authors use a CEGIS-based approach for synthesizing continuous-time
switching controllers that guarantee \emph{reach-while-stay} properties of
closed-loop systems, i.e., properties that specify a set of goal states and
safe states (constrained reachability).  This solution is based on
synthesizing control Lyapunov functions for switched systems that yield
switching controllers with a guaranteed minimum dwell time in each mode. 
However, both approaches are unsuitable for the kind of control we seek to
synthesize.

The work in~\cite{hscc-paper} synthesizes stabilizing controllers for
continuous plants by exploiting bit-accurate verification of
software-implemented digital controllers~\cite{Bessa16} but it is restricted
to closed-loop systems given as transfer function models.

@@

The tool Pessoa~\cite{mazo2010pessoa} synthesizes correct-by-design embedded
control software in a Matlab toolbox.  It is based on the abstraction of a
physical system to an equivalent finite-state machine and on the
verification of reachability properties on it.  Based on this safety
specification, \mbox{Pessoa} can synthesize embedded controller software for
a range of properties.  The embedded controller software can be more
complicated than the state-feedback control we synthesize, and the
properties available cover more detail.  However, relying on state-space
discretization, \mbox{Pessoa} is likely to incur scalability limitations. 
Along this research line, \cite{Anta2010, liu16} study the synthesis of
digital controllers for continuous dynamics, and~\cite{zamani2014} extends
the approach to the recent setup of Network Control Systems.

@@

\noindent\textbf{Discretization Effects --}
%
The classical approach to control synthesis has often disregarded
digitalization effects, whereas more recently modern techniques have focused
on different aspects of discretization, including delayed
response~\cite{Duggirala2015} and Finite Word Length (FWL) semantics, with
the goal either to verify (e.g.,~\cite{daes20161}) or to optimize
(e.g.,~\cite{oudjida2014design}) given implementations.

There are two different problems that arise from FWL semantics.  The first
is the error in the dynamics caused by the inability to represent the exact
state of the physical system, while the second relates to rounding and
saturation errors during computation.  In~\cite{fialho1994stability}, a
stability measure based on the error of the digital dynamics ensures that
the deviation introduced by FWL does not make the digital system unstable. 
A~more recent approach~\cite{DBLP:journals/automatica/WuLCC09} uses the
$\mu$-calculus to directly model the digital controller so that the selected
parameters are stable by design.  The analyses
in~\cite{DBLP:conf/hybrid/RouxJG15, DBLP:conf/hybrid/WangGRJF16} rely on an
invariant computation on the discrete system dynamics using Semi-Definite
Programming (SDP).  While the former uses bounded-input and bounded-output
(BIBO) properties to determine stability, the latter uses Lyapunov-based
quadratic invariants.  In both cases, the SDP solver uses floating-point
arithmetic and soundness is checked by bounding the error.  An alternative
is~\cite{park2016scalable}, where the verification of given control code is
performed against a known model by extracting an LTI model of the code by
symbolic execution: to account for rounding errors, an upper bound is
introduced in the verification phase.  The work
in~\cite{picasso2003stabilization} introduces invariant sets as a mechanism
to bound the quantization error effect on stabilization as an invariant set
that always converges toward the controllable set. 
Similarly, \cite{liberzon2003hybrid} evaluates the quantization error
dynamics and bounds its trajectory to a known region over a finite time
period.  This technique works for both linear and non-linear systems.

@@

We are interested in control systems with an output $\vec{y}$ that is a
linear combination of its states and inputs\footnote{Notice that since we
can only measure $\vec{y}$, the state-space $\vec{x}$ is not fully
observable.}.
%
%, which may restrict the observability of the state-space from the output space.
\begin{equation}
\label{eq:ode2}  
\vec{y}(t)=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t),
\vec{y} \in \mathbb{R}^o,
\mat{C} \in \mathbb{R}^{o \times p},
\mat{D} \in \mathbb{R}^{o \times q}. 
\end{equation}
%
Equations~\eqref{eq:ode1}-\eqref{eq:ode2} are soundly discretized in
time~\cite{middleton1990digital,van1978computing} at constant intervals,
whose duration we refer to as the sample time $T_s$, into the difference
equation
%
\begin{align}
\label{eq:plant}
\left\{
\begin{array}{rl}
\vec{x}_{k+1} =& \mat{A}_d \vec{x}_k+ \mat{B}_d \vec{u}_k\\
\vec{y}_{k} =& \mat{C} \vec{x}_k + \mat{D} \vec{u}_k, 
\end{array}
\right.
\text{ where }
\left\{
\begin{array}{l}
\mat{A}_d=e^{\mat{A}T_s}\\
\mat{B}_d = \int_{t = 0}^{T_s} e^{\mat{A} t} dt\ \mat{B}
\end{array}
\right.,
\end{align} 
%
%where $\mat{A}_d=e^\mat{A}$, $\mat{B}_d = \int_{t = 0}^{T_s} e^{\mat{A} t} dt\ \mat{B}$,
%
$k \in \mathbb N$ is a discrete counter and $\vec{x}_{0}=\vec{x}(0)$ is the
initial state.  In this work we raise the standard assumption that
$\mat{D}=\mat{0}$~\cite{Astrom08}.
%
%Also, since $\vec{y}$ can be directly derived from $\vec{x}$, we will omit discussing it in the following unless strictly necessary.

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Semantics of LTI models}\label{sec:model_semantics}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%
The traces of model \eqref{eq:plant}, starting from an initial set
$X_0\subseteq \mathbb{R}^p$, with inputs taken from the set $U \subseteq
\mathbb{R}^q$, are sequences
$ \vec{x}_0 \xrightarrow{\vec{u}_0} \vec{x}_1 \xrightarrow{\vec{u}_1} \vec{x}_2 \xrightarrow{\vec{u}_2} \ldots $, 
%
where $ \vec{x}_0 \in X_0$ and $\forall k\geq 0, \vec{u}_k \in U \text{ and } \vec{x}_{k+1} = \tau(\vec{x}_k,\vec{u}_k) $, 
where 
%
\begin{equation}\label{equ:reachtraj}
\tau(\vec{x}_k,\vec{u}_k) = %\big\{ 
\mat{A}\vec{x}_k + \mat{B}\vec{u}_k
 %\mid \mat{G}\vec{x}_k \leq \vec{h}\big\}\;. 
\end{equation}
%\note{do we actually need to mention the safety spec here given that it is formally introduced only in Sec~\ref{sec:cof_safety}?}
%
%Here $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe a safe set, within the state-space $\mathbb{R}^p$.
%
We generalize this notation to convex sets of initial conditions and inputs
($X_0$ and $U$, respectively), and write $\tau(X_0,U)$ to denote the set of
states $\{\vec{x} \mid \vec{x} = \tau(\vec{x}_0,\vec{u}),\, \vec{x}_0 \in
X_0,\, \vec{u} \in U \}$ reached from $X_0$ by $\tau$ in one step.
%
We furthermore define the $n$-reach set $\tau^n(X_0,U)$ as the set of states
reached from $X_0$ via $\tau$ in $n$ steps.
%
\begin{align}\label{equ:reachset}
\tau^n(X_0,U) = \{\vec{x}_n \mid & 
\forall k\in [0,n-1] \, \vec{x}_{k+1}=\tau(\vec{x}_{k},\vec{u}_{k}), \,
\vec{u}_{k} \in U, \,   
\vec{x}_0 \in X_0
\} \;. 
\end{align}
%
Since the transformations $\mat{A}$ and $\mat{B}$ are linear and linear
transformations preserve convexity~\cite{boyd2004convex}, the sets $X_n =
\tau^n(X_0,U)$ are also convex.  We define the \emph{$n$-reach tube} as
%
\begin{equation}\label{equ:reachtube}
\hat{X}_n=\hat{\tau}^n(X_0,U)=\bigcup_{k\in[0,n]} \tau^k(X_0,U)
\end{equation}
%
as the union of $k$-reach sets over $n$ iterations. 
%
Finally, the set $\hat{X} =\bigcup_{n\geq 0} \tau^n(X_0,U)$, which we refer
to as the reach space of the system, generalizes the previous notions over
an unbounded time horizon (transitive closure).

@@

A closed-loop digital control interconnection with an observer is
illustrated in Figure~\ref{fig:observersystem}.  The linear feedback
controller (bottom left) is characterized by a gain matrix $\mat{K}_{\langle
C \rangle}$ that influences the closed-loop dynamics, and is driven by the
observer state $\hat{\vec{x}}$.  Moreover, $\vec{y}_e=\hat{\vec{y}} -
\vec{y}$ is the error on the output estimator,
%
%  defined as the difference between the real output $\vec{y}_k$ and estimated one $\hat{\vec{y}_k}$, 
%
which is amplified by a gain matrix $\mat{L}_{\langle C \rangle}$.  Note
that the controller and the observer are implemented in a discrete domain
(quantities within this domain are identified by the index $\langle C
\rangle$), whereas the plant evolves in a continuous domain.  We~further
model the effects of using a discrete controller as the noise signals
$\nu_1$(ADC quantization), $\nu_2$ (DAC quantization), $\nu_1'$ (rounding of
the observer state error), $\nu_2'$ (rounding of the observer state), and
$\nu_t$ (output delay of the digital system) Lastly, $k_r$ is a reference
gain required by the model.  Our goal in this work is to synthesize both
$\mat{K}_{\langle C \rangle}$ and $\mat{L}_{\langle C \rangle}$, so that
specific closed-loop properties of the overall interconnection are
satisfied.

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\section{Acceleration of continuous-time models}\label{sec:continuous_time_accel}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

The time discretization in Sec.~\ref{sec:model} only allows us to evaluate
the state of the plant at integer multiples of the sample time.  However, in
order to synthesize $\mat{K}=\mat{K}_{\langle C \rangle}$ and
$\mat{L}=\mat{L}_{\langle C \rangle}$, we must be able to ensure that
certain properties of the plant hold at any given real time.  In order to do
so, we accelerate the dynamics of the plant according to the model semantics
in Sec.~\ref{sec:model_semantics} and compute the state $\vec{x}_T$ at a
given time $T$,
%
% namely we find an equation for $\tau^n(\vec{x},\vec{u})$ given $\tau(\vec{x},\vec{u})$. 
%
namely the state of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time
$t=T$.

%\note{ [add?] In this section we work with fully observed models, namely where $y = x$. } 
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Accelerated\! dynamics\! for\! continuous-time\! models\! without\! inputs}\label{sec:cont_acc_no_inputs}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%
Given a specific time $T$, the state $\vec{x}(t=T)$ can be expressed as a
linear function of $\vec{x}_0$ as shown below.  We assume that the system
has no inputs, so its dynamics are described by
$\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$.

\begin{lemma}
The solution to the differential equation $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$, where $\mat{A}=\mat{S}\mat{J}\mat{S}^{-1}$, and $\mat{S}$ are the generalized eigenvectors of $\mat{A}$,
evaluated at time $T$ is
%
\begin{align}
 \vec{x}_T&=\vec{x}(t=T)=\mat{A}_{T}\vec{x} (0)\\
 \mat{A}_{T}&= \mat{S}
 \left [ \begin{array}{cccc}
 e^{T\lambda_1}  & s_1\frac{T^{1}e^{T\lambda_i}}{(2)!} & \hdots  & s_i\frac{T^{p-1}e^{T\lambda_i}}{(p-i)!} \\
0 & e^{T\lambda_i}  & s_i\frac{T^{j-i}e^{T\lambda_i}}{(j-i)!} & \vdots \\
\vdots & & \ddots & \vdots \\
0 & \cdots & 0  &e^{T\lambda_i} \\
\end{array} \right ]
 \mat{S}^{-1}
 \label{eq:continuous_tube_dyn2}\\
 &\text{where } s_i=\left\{\begin{array}{cc}1&\quad gm(\lambda_i)>1\\0&\quad gm(\lambda_i)=1\end{array}\right.,\nonumber
\end{align}
%
$\lambda_i$ are the eigenvalues of $\mat{A}$, and $gm(\lambda_i)$ is the geometric multiplicity of $\lambda_i$.
%
%$\vec{x}_T$ is the state of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time t=T. 
If $\mat{J}$ is diagonal and there exists a discrete dynamics matrix for a sampling time $T_d$ such that $A_d=e^{\mat{A} T_d}$, then $\vec{x}_T=A_d^{\frac{T}{T_d}}\vec{x}_0 \text{ with } T \in \mathbb{R}^+_0$ is the state of the system at time $t=T$. 
\end{lemma}

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Accelerated dynamics for models with parametric inputs}\label{sec:real_discrete_param_inputs}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

Next, we are interested in studying models with parametric inputs, which are
described by $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$, where
$\forall t\geq 0, \vec{u}(t)=\vec{u}$.  For such models, a state $\vec{x}_T$
is computed next.
%
\begin{lemma}
%
The state of the model
$\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$ at time $T$ given a
parametric input $\vec{u}$ is
%
\begin{align}
\vec{x}_T&=\vec{x}(t=T)=\mat{A}_T\vec{x}_0 + \mat{B}_T\vec{u} \text{ where } \mat{B}_T=\mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B} 
\label{eq:continuous_tube_param}
\end{align}
%
where $\forall t \leq T,\ \vec{u}(t)=\vec{u}$, and $\mat{A}_T$ as in
\eqref{eq:continuous_tube_dyn2} \footnote{When $\mat{A}$ is not invertible,
we use $\left[\begin{array}{cc}\mat{A}_T&\mat{B}_T\\
0&\mat{I}\end{array}\right]=e^{\left[\begin{array}{cc}\mat{A}&\mat{B}\\0&0\end{array}\right]}$}. 
If $\mat{J}$ is diagonal and given a sampling time $T_d$ such that
$A_d=e^{\mat{A} T_d}$, then
$\vec{x}_T=A_d^{\frac{T}{T_d}}\vec{x}_0+(\mat{I}-\mat{A}_d^{\frac{T}{T_d}}\mat{B}_d\vec{u})
\text{ with } T \in \mathbb{R}_0^+$ (where $\mat{A}_d^{\frac{T}{T_d}} =
e^{log(\mat{A}_d) \frac{T}{T_d}} = e^{\mat{A} T}$) is the state of the
system at time $T$.
%
\end{lemma}
